---
title: "Practical_324"
author: "Robert Atkinson"
date: "07/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## load libraries
```{r load libraries, echo=TRUE}

library(tidyverse)
library(magrittr)
library(psych)
library(stargazer)
library(zoo)
library(car)
library(lmtest)
library(lm.beta)
```


## load data
```{r}

leicester_2011OAC <- readr::read_csv("2011_OAC_Raw_uVariables_Leicester.csv")

```




# Exercise 324.1

## Question 324.1.1: 

Create a model having as outcome variable the presence of people using private transport for commuting to work, and using a stepwise “both” approach, having all the variables created for the example above and related to the presence of people working in different industry sectors (perc_u141 to perc_u158) as scope.

```{r}
# Select and
# normalise variables
leicester_2011OAC_transp_q1 <-
leicester_2011OAC %>%
dplyr::select(
OA11CD,
Total_Pop_No_NI_Students_16_to_74, Total_Employment_16_to_74,
u121, u141:u158
) %>%
# percentage method of travel
dplyr::mutate(
u121 = (u121 / Total_Pop_No_NI_Students_16_to_74) * 100
) %>%
# percentage across industry sector columns
dplyr::mutate(
dplyr::across(
u141:u158,
function(x){ (x / Total_Employment_16_to_74) * 100 }
)
) %>%
# rename columns
dplyr::rename_with(
function(x){ paste0("perc_", x) },
c(u121, u141:u158)
)
```

```{r echo=TRUE, warning=FALSE}

library(MASS)
# Selected variables
# perc_u120: Method of Travel to Work, Private Transport
# perc_u142: Industry Sector, Mining and quarrying
# perc_u144: Industry Sector, Electricity, gas, steam and air conditioning ...
# perc_u146: Industry Sector, Construction
# perc_u149: Industry Sector, Accommodation and food service activities
# Create model
commuting_model_q1 <-
leicester_2011OAC_transp_q1 %$%
stepAIC(
  object = 
lm(perc_u121 ~ 1),
scope =
perc_u121 ~
  perc_u141 + perc_u142 + perc_u143 + perc_u144 + perc_u145 + perc_u146 + perc_u147 + perc_u148 + perc_u149 +
  perc_u150 + perc_u151 + perc_u152 + perc_u153 + perc_u154 + perc_u155 + perc_u156 + perc_u157 + perc_u158
)
# Print summary
stargazer(commuting_model_q1, type = "text")

```



```{r}
commuting_model_q1 %>%
rstandard() %>%
stats::shapiro.test()
```


Overall, we can say that the delay model computed above is fit (F(13, 955) = 74.83, p < .001), indicating
that the model might account for  49.79% of the people who use private tansport to commute. However the model is not
robust. The residuals are not homoscedastic (Breusch-Pagan test, BP = 27.606, p < .01),
and are not independent (Durbin-Watson test, DW = 1.8471, p < 0.01), but they are normally
distributed (Shapiro-Wilk test, W = 0.99618, p < 0.01).



```{r}
commuting_model_q1 %>%
lmtest::bptest()
```



```{r}
commuting_model_q1 %>%
lmtest::dwtest()
```


## Question 324.1.2: 

Is the presence of people using public transportation to commute to work statistically, linearly related to mean age (u020)?


```{r}

# Select and normalise variables
leicester_2011OAC_transp_q2 <-
leicester_2011OAC %>%
dplyr::select(
OA11CD,
Total_Pop_No_NI_Students_16_to_74, Total_Employment_16_to_74,
u020, u120
) %>%
# percentage method of travel
dplyr::mutate(
perc_u020 = (u020/ Total_Pop_No_NI_Students_16_to_74) * 100,  
perc_u120 = (u120/ Total_Pop_No_NI_Students_16_to_74) * 100
) 

```




#### Check assumptions for Pearson's


```{r}

leicester_2011OAC_transp_q2 %$% perc_u020 %>%
stats::shapiro.test()

leicester_2011OAC_transp_q2 %>%
  ggplot2::ggplot(mapping = aes(x = perc_u020)) +
  ggplot2::geom_histogram(bins = 30) + 
  ggplot2::theme_gray()
```

Not normally distributed. 

```{r perc_120 test for normal distribution}
leicester_2011OAC_transp_q2 %$% perc_u120 %>%
stats::shapiro.test()

leicester_2011OAC_transp_q2 %>%
  ggplot2::ggplot(mapping = aes(x = perc_u120)) +
  ggplot2::geom_histogram(bins = 30) + 
  ggplot2::theme_gray()
```

Is normally distributed. 



```{r}

log_mean_age <- leicester_2011OAC_transp_q2 %>%
mutate(log10_u020 = log10(perc_u020))

  
log_public_commute <- leicester_2011OAC_transp_q2 %>%
mutate(log10_u120 = log10(perc_u120))


log_mean_age %$% log10_u020 %>%
stats::shapiro.test()

log_public_commute %$% perc_u120 %>%
stats::shapiro.test()
```
Both variables are not normaly distributed after logarithmic transformation. 


Pearson's can't be used as both variables need to be norally distributed. 

#### Check Spearson's assumptions

```{r}
ties_check_mean_age <-
leicester_2011OAC_transp_q2 %>%
dplyr::count(perc_u020) %>%
dplyr::filter(n > 1) %>%
dplyr::count(n()) %>%
dplyr::pull(n)

ties_check_mean_age

```

```{r}
ties_check_public_commute <-
leicester_2011OAC_transp_q2 %>%
dplyr::count(perc_u120) %>%
dplyr::filter(n > 1) %>%
dplyr::count(n()) %>%
dplyr::pull(n)

ties_check_public_commute
```
There are too many ties in the perc_u120 data. Spearson's cannot be used. 


#### Kendall's Tau


```{r}
 

kendall_mean_age_public_commute <-
leicester_2011OAC_transp_q2 %$%
stats::cor.test(
perc_u020, perc_u120,
method = "kendall"
)

kendall_mean_age_public_commute
```


Mean age has a very small but significant linear relationship with people using public transportation to commute to work  (Tau = 0.08,  p < 0.001).



## Question 324.1.3:

Is the presence of people using public transportation to commute to work statistically, linearly related to (a subset of) the age structure categories (u007 to u019)?


```{r results='asis', echo = TRUE, warning = FALSE}
leicester_2011OAC_transp_q3 <-
leicester_2011OAC %>%
dplyr::select(
OA11CD,
Total_Population, Total_Pop_No_NI_Students_16_to_74, Total_Employment_16_to_74,
u007:u019, u120
) %>%
# percentage method of travel
dplyr::mutate(
perc_u120 = (u120/ Total_Pop_No_NI_Students_16_to_74) * 100
) %>%
#percentage across industry sector columns
dplyr::mutate(
dplyr::across(
u007:u019,
function(x){ (x / Total_Population) * 100 }
)
) %>%

# rename columns
dplyr::rename_with(
function(x){ paste0("perc_", x) },
c(u007:u019)
)

leicester_2011OAC_transp_q3 %>%
dplyr::select(perc_u120, perc_u007, perc_u008, perc_u009, perc_u010, perc_u011, perc_u012, perc_u013, perc_u014, perc_u015, perc_u016, perc_u017, perc_u018, perc_u019) %>%
psych::pairs.panels(method = "kendal", stars = TRUE)
 

commuting_model_q3 <-
leicester_2011OAC_transp_q3 %$%
stats::lm(
perc_u120 ~ perc_u013) 


# Print summary
stargazer(commuting_model_q3, type = "text")


```

Age 25 - 29 (u012) has a small but significant linear relationship with people using public transportation to commute to work  (Tau = 0.11,  p < 0.001).

Age 30 - 44 (u013) also has a small but significant linear relationship with people using public transportation to commute to work  (Tau = 0.14,  p < 0.001).



